### 2024-09-24 manuscript.tex
Improving the methods section

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between 2010
and 2021 (Fig. \ref{fig:map}, Table \ref{tab:voyages} and
\ref{tab:voyage-references}). Altogether, the voyages and the station comprised
2421 days of data south of 40°S, but the availability of ceilometer data was
smaller due to gaps in measurements. Missing days in the ceilometer data were
HMNZSW16 (7 days): 24--27 November, 10 December, 16--17 December 2016; MARCUS
(3 days): 8, 10 November, 10 December; MICRE (9 days): 7--8, 29 June, 5, 16
July, 15 August, 17 October 2016, 11 February, 21 March 2017; TAN1502 (1 day):
24 January.

These data sources contained ceilometer observations with the Vaisala CL51 and
CL31 operating at 910 nm and the Lufft CHM-15k operating at 1064 nm, described
in detail below (Sections \ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is
a low-power near-infrared vertically pointing lidar with the purpose of
measuring the cloud base, but it also measures the full vertical structure of
clouds as long as the laser signal is not attenuated by thick clouds.

Apart from lidar observations, radiosondes were launched on weather balloons at
regular synoptic times on the RV \emph{Polarstern}, MARCUS, NBP17024, TAN1702,
and TAN1802 voyages, measuring pressure, temperature, relative humidity, and
global navigation satellite system coordinates. Derived thermodynamic (virtual
potential temperature, lifting condensation level, etc.) and dynamic quantities
(wind speed and direction) for the measured vertical profiles were calculated
with rstool \citep{rstool}. Surface meteorological quantities were measured
continuously by an onboard automatic weather station or individual instruments.

@@

\subsection{ALCF}

The Automatic Lidar and Ceilometer Framework (ALCF) is a ground-based lidar
simulator. It performs radiative transfer calculations to derive equivalent
lidar backscattered radiation in an atmospheric model, which can then be
compared with observed lidar backscatter \citep{kuma2021}. For this purpose, it
takes atmospheric fields of cloud fraction, liquid and ice mass mixing ratio,
temperature, and pressure fields as an input and can be run offline (on the
model output rather than inside the model code). The lidar simulator in the
ALCF is based on the instrument simulator Cloud Feedback Model Intercomparison
Project (CFMIP) Observation Simulator Package (COSP) \citep{bodas-salcedo2011}.
After the backscattered radiation is calculated, a cloud mask, cloud occurrence
by height, and the total cloud fraction are determined. ALCF has already been
used by several research teams for model and reanalysis evaluation
\citep{kuma2020,kremser2021,guyot2022,pei2023,whitehead2023,mcdonald2024}.
Absolute calibration of the observed backscatter was performed by comparing the
measured clear-sky molecular backscatter statistically with simulated clear-sky
molecular backscatter.

The Vaisala CL51 (Fig. \ref{fig:map}d) and CL31 are ceilometers operating at a
near-infrared wavelength of 910 nm. The maximum range is 15.4 km (CL51), 7.7 km
(CL31 with 5 m vertical resolution), and 7.550 km (CL31 with 10 m vertical
resolution).  The vertical resolution is 10 m (5 m configurable). The
wavelength is affected by water vapour absorption of about 20\% in the
mid-latitudes \citep{wiegner2015,wiegner2019}.  The instrument data files
containing uncalibrated backscatter were first converted to NetCDF with cl2nc
(\url{https://github.com/peterkuma/cl2nc}) and then processed with the ALCF to
produce a cloud mask and cloud occurrence by height and the total cloud
fraction. CL51 can also be configured to emulate CL31.

For the MARCUS voyage, Vaisala CL31 was made available by the Atmospheric
Radiation Measurement (ARM) processed to a maximum range of 7.545 km, a
vertical resolution of 30 m, and 252 vertical levels.

The Lufft CHM 15k (Fig. \ref{fig:map}c) is a ceilometer operating at a
near-infrared wavelength of 1064 nm. The maximum range is 15.4 km, the vertical
resolution is 5 m in the near range (up to 150 m) and 15 m above, the sampling
rate is 2 s, and the number of vertical levels is 1024.  NetCDF files
containing uncalibrated backscatter produced by the instrument were processed
with the ALCF to produce a cloud mask and cloud occurrence by height and the
total cloud fraction.

% TODO: Check sampling rate.

@@

\begin{table}
% TODO: Add sampling intervals.
\caption{
List of voyages and stations. Start, end, and days (UTC; inclusive)
indicate the time period when the vessel was south of 40°S. Abbreviations:
ceilometer (ceil.), Australia (AU), New Zealand (NZ), South America (SA),
Atlantic Ocean (AO) and Africa (AF). The number of days is rounded to the
nearest integer. CL51/31 indicates CL51 configured to emulate CL31.
}
\label{tab:voyages}
\small
\begin{tabular}{llllllr}
\textbf{Name} & \textbf{Vessel} & \textbf{Ceil.} & \textbf{Region} & \textbf{Start} & \textbf{End} & \textbf{Days}\\
\hline
AA15-16  & RSV \emph{Aurora Australis}   & CL51    & AU       & 2015-10-22 & 2016-02-22 & 124 \\
HMNZSW16 & HMNZS \emph{Wellington}       & CHM 15k & NZ       & 2016-11-23 & 2016-12-19 & 27 \\
MARCUS   & RSV \emph{Aurora Australis}   & CL31    & AU       & 2017-10-29 & 2018-03-26 & 149 \\
MICRE    & [Macquarie Is. station]       & CL51    & AU/NZ    & 2016-04-03 & 2018-03-14 & 710 \\
NBP1704  & RV \emph{Nathaniel B. Palmer} & CHM 15k & NZ       & 2017-04-14 & 2017-06-08 & 55 \\
PS77/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2010-12-01 & 2011-02-04 & 65 \\
PS77/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-02-07 & 2011-04-14 & 66 \\
PS79/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-12-06 & 2012-01-02 & 27 \\
PS79/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-01-10 & 2012-03-10 & 61 \\
PS79/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-03-14 & 2012-04-08 & 26 \\
PS81/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-12-02 & 2013-01-18 & 47 \\
PS81/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-01-22 & 2013-03-17 & 55 \\
PS81/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-03-18 & 2013-04-16 & 30 \\
PS81/5   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-04-20 & 2013-05-23 & 33 \\
PS81/6   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-06-10 & 2013-08-12 & 63 \\
PS81/7   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-08-15 & 2013-10-14 & 60 \\
PS81/8   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-11-12 & 2013-12-14 & 31 \\
PS81/9   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-12-21 & 2014-03-02 & 71 \\
PS89     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2014-12-05 & 2015-01-30 & 56 \\
PS96     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2015-12-08 & 2016-02-14 & 68 \\
PS97     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-02-15 & 2016-04-06 & 52 \\
PS103    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-12-18 & 2017-02-02 & 46 \\
PS104    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2017-02-08 & 2017-03-18 & 39 \\
PS111    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-01-21 & 2018-03-14 & 52 \\
PS112    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-03-18 & 2018-05-05 & 49 \\
PS117    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-12-18 & 2019-02-07 & 51 \\
PS118    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2019-02-18 & 2019-04-08 & 50 \\
PS123    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-01-10 & 2021-01-31 & 21 \\
PS124    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-02-03 & 2021-03-30 & 55 \\
TAN1502  & RV \emph{Tangaroa}            & CL51/31 & NZ       & 2015-01-20 & 2015-03-12 & 51 \\
TAN1702  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2017-03-09 & 2017-03-31 & 23 \\
TAN1802  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2018-02-07 & 2018-03-20 & 41 \\
\hline
\textbf{Total} &                         &         &          &            &            & \textbf{2421}\\
\end{tabular}
\normalsize
\end{table}

>>>>

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between 2010
and 2021. Fig. \ref{fig:map} shows a map of the voyages, Table \ref{tab:voyages}
list the voyages and stations, and \ref{tab:voyage-references} lists references
where available. Altogether, the voyages and station dataset comprised
2421 days of data south of 40°S, but the availability of ceilometer data was
slightly smaller due to gaps in measurements.

Missing days in the ceilometer data were HMNZSW16 (7 days): 24--27 November, 10
December, 16--17 December 2016; MARCUS (3 days): 8, 10 November, 10 December;
MICRE (9 days): 7--8, 29 June, 5, 16 July, 15 August, 17 October 2016, 11
February, 21 March 2017; TAN1502 (1 day): 24 January.

The data sources contained ceilometer observations captured by the Vaisala CL51
and CL31 operating at a wavelenght of 910 nm and the Lufft CHM-15k operating at
1064 nm, described in detail below (Sections \ref{sec:cl51} and
\ref{sec:chm15k}). A ceilometer is a low-power near-infrared vertically
pointing lidar with the purpose of measuring the cloud base, but it also
measures the full vertical structure of clouds as long as the laser signal is
not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.

Apart from lidar observations, radiosondes were launched on weather balloons at
regular synoptic times on the RV \emph{Polarstern}, MARCUS, NBP17024, TAN1702,
and TAN1802 voyages, measuring pressure, temperature, relative humidity, and
global navigation satellite system coordinates. Derived thermodynamic (virtual
potential temperature, lifting condensation level, etc.) and dynamic physical
quantities (wind speed and direction) for the measured vertical profiles were
calculated with rstool \citep{rstool}. Surface meteorological quantities were
measured continuously by an onboard automatic weather station or individual
instruments.

The Vaisala CL51 (photo in Fig. \ref{fig:map}d) and CL31 are ceilometers
operating at a near-infrared wavelength of 910 nm. The maximum range is 15.4 km
(CL51), 7.7 km (CL31 with 5 m vertical resolution), and 7.550 km (CL31 with 10
m vertical resolution). The vertical resolution is 10 m (5 m configurable), and
the sampling (temporal) resolution is 6 s (AA15‐16, MARCUS, and MICRE), 36 s (RV
\emph{Polarstern}), and about 2.37 s (TAN1502). From the MARCUS voyage, unlike raw
instrument data available for the other voyages and stations, the Vaisala CL31
data were made available by the Atmospheric Radiation Measurement (ARM)
processed to a maximum range of 7.545 km, a vertical resolution of 30 m, and
sampling resolution of 16 s. The wavelength of 910 nm is affected by water
vapour absorption of about 20\% in the mid-latitudes
\citep{wiegner2015,wiegner2019}.  The instrument data files containing raw
uncalibrated backscatter were first converted to NetCDF with cl2nc
(\url{https://github.com/peterkuma/cl2nc}) and then processed with the ALCF
(Section \ref{sec:alcf}) to produce absolutely calibrated attenuated volume
backscattering coefficinet (AVBC), cloud mask, cloud occurrence by height and
the total cloud fraction. CL51 can also be configured to emulate CL31. The
Vaisala CL51 and CL31 instruments were used on most of the voyages and stations
analysed here. Fig. \ref{fig:example}a shows an example of derived from the
CL51 instrument.

The Lufft CHM 15k (photo in Fig. \ref{fig:map}c) is a ceilometer operating at a
near-infrared wavelength of 1064 nm. The maximum range is 15.4 km, the vertical
resolution is 5 m in the near range (up to 150 m) and 15 m above, the sampling
(temporal) resolution is 2 s, and the number of vertical levels is 1024.
NetCDF files containing uncalibrated backscatter produced by the instrument
were processed with the ALCF (Section \ref{sec:alcf}) to produce AVBC, cloud
mask, cloud occurrence by height and the total cloud fraction. The CHM 15k was
used on a minority of voyages (HMNZSW16, TAN1702, TAN1802, and NBP1704).

\subsection{ALCF}
\label{sec:alcf}

The Automatic Lidar and Ceilometer Framework (ALCF) is a ground-based lidar
simulator and a tool for processing observed lidar data, supporting various
instruments and models. It performs radiative transfer calculations to derive
equivalent lidar AVBC in an atmospheric model, which can then be compared with
observed AVBC \citep{kuma2021}. For this purpose, it takes atmospheric fields
of cloud fraction, liquid and ice mass mixing ratio, temperature, and pressure
fields as an input and can be run offline (on the model output rather than
inside the model code). The lidar simulator in the ALCF is based on the
instrument simulator Cloud Feedback Model Intercomparison Project (CFMIP)
Observation Simulator Package (COSP) \citep{bodas-salcedo2011}.  After AVBC is
calculated, a cloud mask, cloud occurrence by height, and the total cloud
fraction are determined. The ALCF has already been used by several research
teams for model and reanalysis evaluation
\citep{kuma2020,kremser2021,guyot2022,pei2023,whitehead2023,mcdonald2024}.
Absolute calibration of the observed backscatter was performed by comparing the
measured clear-sky molecular backscatter statistically with simulated clear-sky
molecular backscatter. AVBC is resampled to 5 min temporal resolution and 50 m
vertical resolution to increase signal-to-noise ratio, while having enough
resolution to detect small-scale cloud variability. Noise standard deviation is
calculated from AVBC at the highest range, where no clouds are expected.  A
cloud mask is calculated from AVBC using a fixed threshold of $\mathrm{2\times
10^{-6} m^{-1}sr^{-1}}$ after subtracting 5 standard deviations of range-scale
noise. Fig. \ref{fig:example}b shows an example of simulated Vaisala CL51
backscatter from ERA5 data, corresponding to a day of measurements by the
instrument on the PS81/3 voyage.

@@

@@

\begin{table}
% TODO: Add sampling intervals.
\caption{
List of voyages and stations. Start, end, and days (UTC; inclusive)
indicate the time period when the vessel was south of 40°S. Abbreviations:
ceilometer (ceil.), Australia (AU), New Zealand (NZ), South America (SA),
Atlantic Ocean (AO) and Africa (AF). The number of days is rounded to the
nearest integer. CL51/31 indicates CL51 configured to emulate CL31.
}
\label{tab:voyages}
\small
\begin{tabular}{llllllr}
\textbf{Name} & \textbf{Vessel or station} & \textbf{Ceil.} & \textbf{Region} & \textbf{Start} & \textbf{End} & \textbf{Days}\\
\hline
AA15-16  & RSV \emph{Aurora Australis}   & CL51    & AU       & 2015-10-22 & 2016-02-22 & 124 \\
HMNZSW16 & HMNZS \emph{Wellington}       & CHM 15k & NZ       & 2016-11-23 & 2016-12-19 & 27 \\
MARCUS   & RSV \emph{Aurora Australis}   & CL31    & AU       & 2017-10-29 & 2018-03-26 & 149 \\
MICRE    & Macquarie Is. station         & CL51    & AU/NZ    & 2016-04-03 & 2018-03-14 & 710 \\
NBP1704  & RV \emph{Nathaniel B. Palmer} & CHM 15k & NZ       & 2017-04-14 & 2017-06-08 & 55 \\
PS77/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2010-12-01 & 2011-02-04 & 65 \\
PS77/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-02-07 & 2011-04-14 & 66 \\
PS79/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-12-06 & 2012-01-02 & 27 \\
PS79/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-01-10 & 2012-03-10 & 61 \\
PS79/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-03-14 & 2012-04-08 & 26 \\
PS81/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-12-02 & 2013-01-18 & 47 \\
PS81/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-01-22 & 2013-03-17 & 55 \\
PS81/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-03-18 & 2013-04-16 & 30 \\
PS81/5   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-04-20 & 2013-05-23 & 33 \\
PS81/6   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-06-10 & 2013-08-12 & 63 \\
PS81/7   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-08-15 & 2013-10-14 & 60 \\
PS81/8   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-11-12 & 2013-12-14 & 31 \\
PS81/9   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-12-21 & 2014-03-02 & 71 \\
PS89     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2014-12-05 & 2015-01-30 & 56 \\
PS96     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2015-12-08 & 2016-02-14 & 68 \\
PS97     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-02-15 & 2016-04-06 & 52 \\
PS103    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-12-18 & 2017-02-02 & 46 \\
PS104    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2017-02-08 & 2017-03-18 & 39 \\
PS111    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-01-21 & 2018-03-14 & 52 \\
PS112    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-03-18 & 2018-05-05 & 49 \\
PS117    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-12-18 & 2019-02-07 & 51 \\
PS118    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2019-02-18 & 2019-04-08 & 50 \\
PS123    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-01-10 & 2021-01-31 & 21 \\
PS124    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-02-03 & 2021-03-30 & 55 \\
TAN1502  & RV \emph{Tangaroa}            & CL51/31 & NZ       & 2015-01-20 & 2015-03-12 & 51 \\
TAN1702  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2017-03-09 & 2017-03-31 & 23 \\
TAN1802  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2018-02-07 & 2018-03-20 & 41 \\
\hline
\textbf{Total} &                         &         &          &            &            & \textbf{2421}\\
\end{tabular}
\normalsize
\end{table}
